---
title: "Time Series Analysis - Bitcoin Competition"
author: "Loan Huynh - s3655461"
output:
  word_document: default
  html_notebook: default
---

# Introduction

The aim of this project is trying to find the best model for forecasting Bit Coin price in next 10 days.</br>
Dataset is collected from https://coinmarketcap.com/ website between 27 April 2013 to 24 Feb 2019.


# Data Exploration

_*Loading data and library*_

```{r message=FALSE, warning=FALSE}
# Load libraries
library(TSA)
library(fUnitRoots)
library(forecast)
library(CombMSC)
library(lmtest)
library(fGarch)
library(rugarch)
library(tseries)
source("residual.analysis.R")
source("TSHandy.r")
source("MASE.R")

# This function is used to fiting model
modelfit <- function(garchOrder, armaOrder, data){
  model <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = garchOrder), 
                   mean.model = list(armaOrder = armaOrder, include.mean = FALSE), 
                   distribution.model = "norm")
  model.fit <- ugarchfit(spec=model,data=data, out.sample = 100)
  return(model.fit)
}


# sortAIC_garch function is sort Garchmodels based on AIC values
sortAIC_garch <- function(list_GarchOrder, resdata)
{
  rs <- data.frame(matrix(ncol = 2))
  colnames(rs) <- c("GarchOrder", "AIC")
  for (i in 1: length(list_GarchOrder))
  {
    garchOrder <- sapply(list_GarchOrder,function(x) unlist(x))[,i]
    model <- garch(resdata, order = garchOrder)
    aic <- AIC(model)
    rs[i,] <- list(toString(garchOrder), aic)
  }
  return(rs[order(rs$AIC),])
}


# This function transform fitted residuals from forcasting ARIMA+GARCH model into fitted BitCoin Price
fitted.values <- function(forecastResult, transformed_series)
{
  fitted.res <- forecastResult@forecast$seriesFor
  invdiff <- diffinv(fitted.res)
  latest_value <- transformed_series[length(transformed_series)]
  fitted.value.list <- c()
  for(i in 2:length(invdiff))
  {
    trans.fitted.value <- latest_value + invdiff[i]
    fitted.value.list <- c(fitted.value.list, exp(trans.fitted.value))
    latest_value <- trans.fitted.value
  }
  return(fitted.value.list)
}


#This function create model and return MASE results for 10 forecasts
checkMASE <- function(model, data_model, data.trans, observed)
{
 # model <- modelfit(garchOrder, armaOrder, data_model)
  forcRes = ugarchforecast(model, data = data_model, n.ahead = 10)
  predicted_result <- fitted.values(forcRes, data.trans)
  rs <- MASE(observed,predicted_result)$MASE
  return(rs)
}

# Load dataset
bitcoin <- read.csv("Bitcoin_Historical_Price.csv")
bitcoin_ts <- data.frame(bitcoin$Close)
rownames(bitcoin_ts) <- bitcoin$Date
bitcoin_ts <- ts(as.vector(bitcoin_ts), start = c(2013,27,4), frequency = 365) # Convert Bitcoin price data into Time Series Object
plot(bitcoin_ts, type= 'o', main="Bit Coin Price Series from 27/04/2013 - 24/02/2019")
```

Bit Coint series plot illustrates that there is no seasonality. This series fluctuates from 2013 to 2017 and then go up signficantly. It reaches a peak at nearly 20000 USD for some days in the end of 2017. After that it falls considerably in 2018 and 2019. It has succeeding points and changes of variance.  

```{r message=FALSE, warning=FALSE}
# Scatter plot
plot(y = bitcoin_ts, x = zlag(bitcoin_ts), main = "Scatter Plot of Bit Coin within Previous Day")
```

Scatter plot shows a strong relationship between Bit Coint Price and its lag.

```{r}
acf(bitcoin_ts, na.action = na.pass, main ="ACF plot of Bit Coin series") # ACF Plot
pacf(bitcoin_ts, na.action = na.pass, main = "PACF plot of Bit Coin Series") # PACF plot
```

ACF plot has decaying pattern. In conclustion, Bit Coin price is a non-stationary series. In the next parts, transformation and difference are applied to transfer non-stationary series into stationary series. </br>

# Data Transformation and Difference
## Data transformation

Log is applied for transforming data because this series skews to the right in its histogram plot and has positive values.

```{r}
# Log Transform
log.bc <- log(bitcoin_ts)
hist(bitcoin_ts, main = "Histogram of Bit Coin Series before transformation", xlab = "Bit Coin Price", cex.main=1)
hist(log.bc, main = "Histogram of Bit Coin Series after transformation", xlab = "Log Transform of Bit Coin", cex.main=1)
```

_*ACF and PACF of transformed data*_

```{r}

acf(log.bc, na.action = na.pass, main = "ACF of transformed Bit Coin", cex.main= 1)
pacf(log.bc, na.action = na.pass,  main = "PACF of transformed Bit Coin", cex.main= 1)
```
After transformation, the series still got decapying pattern in ACF. It means that it is still a non-stationary sereis. ADF Test is applied to validate the stationary of series </br>

```{r}
ar(diff(log.bc)) # Order 31
adfTest(log.bc, lags = 31)
```

The p-value of ADF Test is 0.9232 larger than 0.05. Therefore, Null hypothesis cannot be rejected. It means that series is non-stationary series. </br>

_*Ljung-Box Test*_

```{r}
Box.test(x = log.bc, type = "Ljung-Box")
```

The p-value is less than 5% of significance level. Therefore, null hypothesis of independence is rejected. It means that the series values are dependents or correlated.

## Data Difference

Apply first difference for transformed series. </br>

```{r}
diff.log.bc <- (diff(log.bc, difference =1)) 
plot(diff.log.bc, type= 'o', main = "First difference of Bit Coin series")
abline(h=0)
```
The above series plot looks like moving average model. </br>
Apply ADF Test for this series

```{r}
# ADF Test
ar(diff(diff.log.bc), na.action = na.pass) # Order is 32
adfTest(diff.log.bc, lags = 32)
```
With order of lag is 32, the differenced series becomes stationary </br>


_*Apply McLeod-Li Test*_

```{r}

McLeod.Li.test(y=diff.log.bc, main="McLeod-Li Test Statistics for Bit Coin Returns")
```
It is clearly observed in McLeod-Li Test result that all lags are signficant at 5% significance level. It implies the existence of volatility clustering.

_*Q-Q plot*_

```{r}
qqnorm(diff.log.bc, main="Q-Q plot of 1st Difference Bit Coint Price", cex.main= 1)
qqline(diff.log.bc, col="red")
```

```{r}
shapiro.test(diff.log.bc)
```

Q-Q normal plot shows that there are so many outliers at both tails. The p-value of Shapiro-Wilk test is less than 0.05. It implies that the first difference of bit coin violates normality assumption. </br>

_*ACF and PACF plots*_

```{r}
acf(diff.log.bc , na.action = na.pass, main = "ACF for 1st difference of series", cex.main= 1)
pacf(diff.log.bc , na.action = na.pass, main = "PACF for 1st difference of series", cex.main= 1)
```

There are so many siginficant lags in both ACF and PACF plots. There are no distinct between both plots. </br>
Due to stationarity and changes of variances problems in original series. In this report, ARMA + GARCH model is applied. </br>

# Data Modeling

## ARIMA model

_*EACF*_

```{r}
eacf(diff.log.bc)
```

EACF suggests possibles models like ARIMA(0,1,11),  ARIMA(6,1,7), ARIMA(6,1,8), ARIMA(7,1,7)

_*BIC Table*_

```{r}
bic = armasubsets(y=diff.log.bc,nar=13,nma=13,y.name='test',ar.method='ols')
plot(bic)
```

BIC table suggests ARIMA(6,1,8), ARIMA(12,1,8) </br>

The set of possible ARIMA models is {ARIMA(0,1,11),  ARIMA(6,1,7), ARIMA(6,1,8), ARIMA(7,1,7), ARIMA(12,1,8)}

### Parameter Estimation </br>

Maximum Likelihood method is considered to estimate parameters </br>

```{r message=FALSE, warning=FALSE}
armaList <- list(c(0,1,11), c(6,1,7), c(6,1,8) ,c(7,1,7),c(12,1,8))
armaEstimation <- myCandidate(bitcoin_ts, orderList = armaList, methodType = "ML")
armaEstimation$IC
```

ARIMA(12,1,8) got smallest AIC and BIC result. </br>

_*Coefficient Test*_

Coefficient Test is applied for all possible models based on AIC and BIC result orderly. </br>

_ARIMA(12,1,8)_

```{r}
arima.1218 <- arima(bitcoin_ts, order = c(12,1,8))
coeftest(arima.1218)
```

There are only three coefficients of ar2, ar9 and ma2 not significant at 5%

_ARIMA(6,1,8)_

```{r}
arima.618 <- arima(bitcoin_ts, order = c(6,1,8))
coeftest(arima.618)
```

Most of coefficients for this model are significant, except ma7 and ar2. </br>

_*ARIMA(6,1,7)*_

```{r}
arima.617 <- arima(bitcoin_ts, order = c(6,1,7))
coeftest(arima.617)
```

Most of coefficients for this model are significant, except ma4, ma5 and ma7 

_ARIMA(7,1,7)_

```{r}
arima.717 <- arima(bitcoin_ts, order = c(7,1,7))
coeftest(arima.717)
```

The majority of coefficients in this model is not significant.

_ARIMA(0,1,11)_

```{r}
arima.0111 <- arima(bitcoin_ts, order = c(0,1,11))
coeftest(arima.0111)
```


Coefficients for ma2, ma3, ma4, ma6, ma9 in arima (0,1,11) are not significant.
From coefficients test result and AIC/BIC table, the set of arimal model {ARIMA(6,1,8), ARIMA(12,1,8)} is considered for dianotics check.

### MODEL DIAGNOTICS

_*ARIMA(6,1,8)*_

```{r message=FALSE, warning=FALSE}
residual.analysis(model = arima.618, class = "ARIMA")
```

_*ARIMA(12,1,8)*_

```{r message=FALSE, warning=FALSE}
residual.analysis(model = arima.1218, class = "ARIMA")
```

It is clearly to see that both ARIMA(12,1,8) and ARIMA(6,1,8) failed in Ljung-Box Test and violate normality of residuals. In next part, garch models are selected based on residuals of two these models. 

## GARCH

#### ARIMA(6,1,8)

_*Absolute*_

```{r}
res.618 <- arima.618$residuals
# Apply absolute transform
ab.res.618 <- abs(res.618)
# ACF and PACF plots
acf(ab.res.618,  ci.type="ma",main="ACF plot for absolute residual series ARIMA(6,1,8)", cex.main= 1)
pacf(ab.res.618, main="PACF plot for absolute residual series ARIMA(6,1,8)", cex.main= 1)
```


```{r}
eacf(ab.res.618)
```

No AR/MA is suggested from eacf of absoluated residuales

_*Squared Transform*_

```{r}
sq.res.618 <- res.618^2
acf(sq.res.618, ci.type="ma",main="ACF plot for square residual series ARIMA(6,1,8)", cex.main= 1)
pacf(sq.res.618, main="PACF plot for square residual series ARIMA(6,1,8)", cex.main= 1)
```

```{r}
eacf(sq.res.618)
```

EACF plot suggests ARMA(5,7), ARMA(6,7), ARMA(5,8) => GARCH(7,5), GARCH(7,6), GARCH(8,5) </br>
The possible GARCH models are GARCH(7,5), GARCH(7,6), GARCH(8,5).


```{r}
list_garchOrder <- list(c(7,5), c(7,6), c(8,5))
sortAIC_garch(list_garchOrder, arima.618$residuals)
```

_*ARIMA(12,1,8)*_

```{r}
res.1218 <- arima.1218$residuals
ab.res.1218 <- abs(res.1218)
# ACF and PACF plots
acf(ab.res.1218,  ci.type="ma",main="ACF plot for absolute residual series ARIMA(12,1,8)", cex.main= 1)
pacf(ab.res.1218, main="PACF plot for absolute residual series ARIMA(12,1,8)", cex.main= 1)
```

```{r}
eacf(ab.res.1218)
```

EACF suggest ARMA(5,7), ARMA(6,7), ARMA(5,8) => GARCH(7,5), GARCH(7,6), GARCH(8,5)

```{r}
sq.res.1218 <- res.1218^2
eacf(sq.res.1218)
```
EACF suggests ARMA(1,9), ARMA(4,8), ARMA(5,8), ARMA(4,9) => GARCH(9,1), GARCH(8,4), GARCH(8,5), GARCH(9,4)

Possibles GARCH model: {GARCH(7,5), GARCH(7,6), GARCH(8,5),GARCH(9,1), GARCH(8,4), GARCH(9,4) }

```{r}
list_garchOrder1218 <- list(c(7,5), c(7,6), c(8,5), c(9,1), c(8,4), c(9,4))
sortAIC_garch(list_garchOrder1218, arima.1218$residuals)
```

## ARIMA + GARCH
### ARIMA(6,1,8) + GARCH(7,5)

```{r message=FALSE, warning=FALSE}
model.618_75 <- modelfit(c(7,5), c(6,8), diff.log.bc)
model.618_75
```


### ARIMA(6,1,8) + GARCH(7,6)

```{r message=FALSE, warning=FALSE}
model.618_76 <- modelfit(c(7,6), c(6,8), diff.log.bc)
model.618_76
```


### ARIMA(6,1,8) + GARCH(8,5)

```{r message=FALSE, warning=FALSE}
model.618_85 <- modelfit(c(8,5), c(6,8), diff.log.bc)
model.618_85
```

### ARIMA(12,1,8) + GARCH(7,5)

```{r message=FALSE, warning=FALSE}
model.1218_75 <- modelfit(c(7,5), c(12,8), diff.log.bc)
model.1218_75
```

### ARIMA(12,1,8) + GARCH(7,6)

```{r message=FALSE, warning=FALSE}
model.1218_76 <- modelfit(c(7,6), c(12,8), diff.log.bc)
model.1218_76
```

### ARIMA(12,1,8) + GARCH(9,1)

```{r message=FALSE, warning=FALSE}
model.1218_91 <- modelfit(c(9,1), c(12,8), diff.log.bc)
model.1218_91
```

### ARIMA(12,1,8) + GARCH(8,4)

```{r message=FALSE, warning=FALSE}
model.1218_84 <- modelfit(c(8,4), c(12,8), diff.log.bc)
model.1218_84
```

### ARIMA(12,1,8) + GARCH(9,4)

```{r message=FALSE, warning=FALSE}
model.1218_94 <- modelfit(c(9,4), c(12,8), diff.log.bc)
model.1218_94
```


### ARIMA(12,1,8) + GARCH(8,5)

```{r message=FALSE, warning=FALSE}
model.1218_85<- modelfit(c(8,5), c(12,8), diff.log.bc)
model.1218_85
```

Model ARIMA(6,1,8) + GARCH(7,6) is the best one </br>

# Validation

```{r}
actual_result <- read.csv("Bitcoin_Prices_Forecasts.csv")
observed <- as.numeric(actual_result$Closing.price)
```

_*ARIMA(6,1,8) + GARCH(7,6)*_

```{r}
checkMASE(model.618_76, diff.log.bc, log.bc, observed)
```

_*ARIMA(6,1,8) + GARCH(7,5)*_

```{r}
checkMASE(model.618_75, diff.log.bc, log.bc, observed)
```

_*ARIMA(6,1,8) + GARCH(8,5)*_

```{r}
checkMASE(model.618_75, diff.log.bc, log.bc, observed)
```

_*ARIMA(12,1,8) + GARCH(9,1)*_

```{r message=FALSE, warning=FALSE}
checkMASE(model.1218_91, diff.log.bc, log.bc, observed)
```

_*ARIMA(12,1,8) + GARCH(9,4)*_

```{r message=FALSE, warning=FALSE}
checkMASE(model.1218_94, diff.log.bc, log.bc, observed)
```

_*ARIMA(12,1,8) + GARCH(7,6)*_

```{r}
checkMASE(model.1218_76, diff.log.bc, log.bc, observed)
```



_*ARIMA(12,1,8) + GARCH(8,5)*_

```{r}
checkMASE(model.1218_85, diff.log.bc, log.bc, observed)
```

# CONCLUSION

Model ARIMA(6,1,8) + GARCH (7,6) is the best model. Its predicted values for next 10 days as below

```{r}
forcRes = ugarchforecast(model.618_76, data = data_model, n.ahead = 10, n.roll = 10)
data.frame(fitted.values(forcRes, log.bc))
```


